The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-10-04 paper/ms.tex
model 2 mostly done

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
of stellar $(M_\star, R_\star, L_\star)$ phase-space.
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that appear to be searchable for the 
desired type of planet, and also of the desired stellar type.  These 
``searchable stars'' are the points on the sky for which they think planets 
are observed with 100\% detection efficiency. Accounting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

@@

One immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets of the desired 
type may differ between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval $\Delta M_\star, 
\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
over stellar multiplicity.
Otherwise, one must discuss ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

@@

Errors (1)-(3) in calculating $N_{\rm det}$ above broadly fall under ``radius 
misclassification'', while error (4) falls under ``completeness 
miscalculation''.
Note that in the above list, a planet cannot be both ``of desired type'' and 
``orbiting a star of undesired type''. We ignore this type of error for 
analytic convenience, but \texttt{reintroduce it later during numerics}.

Our approach is step-by-step starting from a very simple scenario.
\begin{itemize}
\item We consider a universe consisting of only the desired types of stars and
planets.  We give analytic expressions for eta(apparent)/eta(true) taking
into account all applicable errors.
%
\item We allow for a distribution of the mass ratio $dn/dq \propto q^\gamma$, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
We give analytic expressions for eta(apparent)/eta(true) taking into 
account errors of type 2.
%NOTE: empirical L(M) maybe OK for numerics -- not for analytics.
%
\item We allow for a nonzero tolerance $\Delta R_p$ in the planet radius, and 
tolerance $\Delta L_\star$ in stellar luminosity, to qualify as "desirable".  
We recalculate eta(apparent)/eta(true).
(LB: I think this has to become numerical to actually evaluate the occurrence
rate correction factor, $X_\Gamma$.)
%
\item We assume the planets have a power-law radius distribution, $d\eta/dR_p
\propto r^\alpha$.  We calculate the apparent $d\eta/dR_p$ taking into account 
errors of type 1-3.  We do so for stars of the desired type within some 
tolerance dL.
%
\item Everything up to now is more-or-less analytic, perhaps supported by 
numerical checking or numerical integration. At the end we summarize the 
radius-dependence of the occurrence rate shifting.
"Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between"

\end{itemize}

We then dig in, w/ numerics, to what happens with different assumed
completeness vs Rp functions.
We note that the the claimed HJ occ rate discrepency btwn RV \&
transit surveys gets much weaker when accounting for binarity.
This is b/c the more believable occ rate papers do not heavily rely on their
completeness calibration -- they know they're complete.
This is a much more understandable regime than the wonky completeness estimate 
dependent zone of eta-Earth for Kepler.

@@

In passing, at the end, we mention how this changes attempts at eta-Earth 
measurements.

@@

Stellar binarity introduces convoluted effects on transit surveys.
To disentangle them, imagine the following idealized survey.

You observe the entire sky for some duration with a detector of known 
area and bandpass. Your detector is photon-noise limited.
You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
You are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. For instance, G2V dwarfs.
You can only detect your desired type of planet when you observe signals 
with ${\rm S/N} > {\rm (S/N)_{min}}$.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux required for detection, $F_{\rm min}$.

You get funding, and perform the S/N-limited survey. 
You observe all the ``points'' on-sky with apparent 
magnitude $m < m_{\rm min}$, or equivalently with energy flux in your bandpass 
greater than some limit $F > F_{\rm min}$.
All planets with ${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are 
not.

You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twin systems.
	The ratio of the number density of binary systems and the number density of 
	all systems in a volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\begin{itemize}
	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have planets of radius $R_p$, with orbital period $P$.
	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
	systems have planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets at 
	equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
	\end{itemize}
\end{itemize}

The assumption of [Pepper et al, 2003] in a very similar context was that the 
observer correctly pre-selects all of the ``searchable stars''.
This simplification would allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

The above approach ignores incompleteness for binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three distinct maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
	of planet is detectable around single stars.
	These are the selected ``searchable, single stars''.
%
\item $d_{{\rm max},d}$: the maximum distance out to which double stars
  are selected (down to the same minimum flux as that defining $d_{{\rm 
  max},s}$). 
  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{1/2} d^{-1}$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
  are in fact binary systems can yield detectable planets (the ratio of 
  volume in which planets can be detected to the actual selected volume).
  For a more realistic sample, $\gamma_R \approx 0.1$, and there is an 
  interesting asymmetry.
  One finds the completeness for planets which orbit the primary star, 
  $f_{d1,c}$ to be
  \begin{equation}
  f_{d1,c} = (1+\gamma_R)^{-3} \approx 3/4.
  \end{equation}
  However, for planets which orbit the secondary star, the completeness is
  \begin{equation}
    f_{d2,c} = (1+\gamma_R^{-1})^{-3} \approx 3/4000.
  \end{equation}

\end{enumerate}

We note in passing that we are considering a SNR-limited survey, which in this 
case is the same as a magnitude-limited one.
The observer might prefer to perform a completeness-limited survey, which in 
this case would be the same as a volume-limited one with limiting distance
$d_{{\rm max},d}^{\rm p}$.
This change would likely improve interpretability.

With the stage set, we ask:

@@

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
Note that these latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~we count the $\approx 25\%$ of multiple 
systems that are higher order multiples as doubles.

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
stated in terms of a limiting magnitude $m_{{\rm min}}$.
In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. Simply as a 
consequence of imposing a magnitude cut, the maximum distance to which binary 
stars will be selected is greater than that of single stars.
The ratio of double to single systems in the observed SNR-limited sample is 
then
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s}
	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
\end{align}
In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
SNR-limited sample.
Since the completeness fraction for binary systems is $1/8$, there are 
$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
planets as single star systems in the system that can yield planets.

@@

The observer who has never heard of binary star systems wants to find the 
occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars of 
radius $R_1$ and luminosity $L_1$.
They detected $N_{\rm det}$ transit signals that appear to be from planets of 
this desired type.
The occurrence rate they then find, {\it i.e.} the number of planets divided 
by number of ``stars'' (really, stellar systems) is
\begin{equation}
\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
\end{equation}

@@

For the mass-luminosity relation, for analytic convenience we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
As a simple check, we fit lines to mass-luminosity data\footnote{
    Data collected by Torres et al. [2009] for dwarfs earlier than 
    spectral class M, and Benedict et al. [2016] for dwarfs later than 
    spectral class M.
    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
    luminosities, we interpolated over E. Mamajek's 
    table
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float. The 
resulting fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L=M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final addition to our model is that we must assume a stellar mass-radius 
relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
Since again, our main purpose is to develop intuition about observational 
biases, we approximate this as 
\begin{equation}
R\propto M.
\end{equation}
\begin{comment}
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
\end{comment}

In passing, we note that while $w_{d2}=1$ was a reasonable assumption in 
Sec.~\ref{sec:model_1} (since secondaries were identical to primaries), in 
binary systems we now have secondaries with different stellar properties from 
primaries.
Recall that our occurrence rate is defined for planets of a given type 
orbiting stars of a given type.
Assume our universe is one in which the primary mass of all stars is 
$M_1=M_\odot$.
If this is the case, we might define our ``stars of interest'' to be 
those with $0.7<M_\star/M_\odot<1.3$.
Then, in a volume-limited sample, 1 in 3 secondaries would be ``of interest'', 
and could contribute to the true occurrence rate.
The general point is that $w_{d2}$ will be less than 1, unless planets of the 
desired type are more common around secondaries of lower mass (as may actually 
be true for small planets around M dwarfs).
Keeping this in mind, we ask the same questions as for Model \#1.

@@

The true occurrence rate is the average number of planets of the desired type 
per star of the desired type.
The number of planets of the desired type is
\begin{equation}
N_{\rm planets} = \Gamma_{t,s}N_s + \Gamma_{t,d}N_d + 
\Gamma_{t,d}w_{d2}f_{{\rm desired}}N_d,
\end{equation}
where as before $N_s$ and $N_d$ are the number of single and double systems, 
and the fraction $f_{{\rm desired}}$ is defined so that $f_{{\rm desired}} 
N_d$ is the number of secondaries that are of the desired type, and as before 
$\Gamma_{t,d}w_{d2}$ is the average number of planets each of these desired 
stars hosts.

Counting the number of desired stars, $N_s$ is the same as in 
Sec.~\ref{sec:model_1}. However
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3
\end{equation}
for
\begin{equation}
d_{\rm max,d} =
d_{\rm max,s}\times (1+\gamma_R)^{1/2},
\end{equation}
where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.

For a given binary system, $q$, $\gamma_R$, and thus $d_{\rm 
	max,d}(\gamma_R)$ are all random variables.
Since $N_d$ is a function of $d_{\rm max,d}$, 
the number of double star systems becomes a random variable, drawn from the 
ensemble of all possible transit surveys.
Applying the chain rule for probability density functions, the distribution 
function for the number of binary systems,
${\rm prob}(N_d)$, can be written 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\label{eq:prob_Nd_raw}
\end{equation}

@@

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
	\end{center}
	\caption{The joint probability distribution of a binary star's position and 
		mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
		M^{3.5}$. Note that the marginalized 
		mass ratio distribution is uniform at any given distance.}
	\label{fig:joint_prob_r_q}
\end{figure}

>>>>

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
of stellar $(M_\star, R_\star, L_\star)$ phase-space.
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that appear to be searchable for the 
desired type of planet, and also of the desired stellar type.  These 
``searchable stars'' are the points on the sky for which they think planets 
are observed with 100\% detection efficiency. Correcting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

@@

An immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets may differ 
between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval $\Delta M_\star, 
\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
over stellar multiplicity.
Otherwise, one must discuss ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

@@

Errors (1)-(3) in calculating $N_{\rm det}$ can be broadly described as 
``radius misclassification'', while error (4) is an error in the assumed 
survey completeness.
%Note that in the above list, a planet cannot be both ``of desired type'' and 
%``orbiting a star of undesired type''. We ignore this type of error for 
%analytic convenience, but \texttt{reintroduce it later during numerics}.

Our approach is step-by-step starting from a very simple scenario.
In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
desired types of stars and planets.  
We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
of the true to apparent occurrence rate, taking into account all applicable 
errors.
In Sec.~\ref{sec:model_2} we allow for a distribution of the mass ratio $dn/dq 
\propto q^\gamma$, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
numerically for the case of planets orbiting Sun-like dwarf stars.
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
In Sec.~\ref{sec:model_3}, 
we assume the planets have a power-law radius distribution, ${\rm 
d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
all relevant errors.
Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between.
One consequence is that the claimed HJ occurrence rate discrepancy between 
radial velocity \& transit surveys gets much weaker when accounting for 
binarity.
In Sec.~\ref{sec:discussion} we summarize these models and discuss their 
utility in interpreting transit surveys.

\begin{comment}
We then dig in, w/ numerics, to what happens with different assumed
completeness vs Rp functions.

This is b/c the more believable occ rate papers do not heavily rely on their
completeness calibration -- they know they're complete.
This is a much more understandable regime than the wonky completeness estimate 
dependent zone of eta-Earth for Kepler.

@@

In passing, at the end, we mention how this changes attempts at eta-Earth 
measurements.
\end{comment}

@@

Consider the following idealized survey.
An astronomer observes the entire sky for some duration with a detector of 
known area and bandpass.
The detector is photon-noise limited.
They are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. %For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
They are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. %For instance, G2V dwarfs.
They detect the desired type of planet when they observe signals 
above some signal-to-noise floor.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux required for detection, $F_{\rm min}$.
Therefore they observe all of the ``points'' on-sky with apparent 
magnitude $m < m_{\rm min}$, or equivalently with energy flux in their 
bandpass greater than the limiting flux $F > F_{\rm min}$.
After detecting all planets with ${\rm S/N} > {\rm (S/N)_{min}}$, they now 
wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.

Assume this astronomer lives in a universe in which the following assumptions 
hold.
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twins.
	The ratio of the number density of binary systems and the number density of 
	all systems in a volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\begin{itemize}
	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have planets of radius $R_p$, with orbital period $P$.
	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
	systems have planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets at 
	equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
	\end{itemize}
\end{itemize}

The assumption of [Pepper et al, 2003] in a very similar context was that the 
observer can correctly pre-select all of the ``searchable stars''.
This simplification would allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

However, any signal-to-noise limited survey of a population that includes 
binaries cannot ignore incompleteness.
%If we set the limiting flux to give 100\% completeness for single stars, we 
%will unwittingly select binaries near that flux limit, for which dilution 
%will push transit signals below the detection threshold.
This is because any given flux limit maps onto three distinct ``maximum 
distances'' for a population of single stars and twin binaries.
First, there is $d_{{\rm max},s}$: the maximum distance out to which the 
desired type of planet is detectable around single stars. Typically this is 
set to be the distance out to which stars in a SNR-limited survey are 
selected, since these are the ``searchable stars''.
Using the same minimum flux as that defining $d_{{\rm max},s}$, double stars 
are then selected out to a larger limiting distance, $d_{{\rm 
max},d}$~--~effectively a Malmquist bias.

However, the distance out to which double stars are {\it selected} is 
different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
be {\it detected} (around the primary of a double star system).
For a population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
L^{1/2} d^{-1}$,
\begin{equation}
d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
(1 + \gamma_R)^{-1} d_{{\rm max},d}.
\end{equation}
For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
are in fact binary systems can yield detectable planets (the ratio of 
volume in which planets can be detected to the actual selected volume).
For a more realistic sample, $\gamma_R \approx 0.1$, and there is an 
interesting asymmetry.
One finds the completeness for planets which orbit the primary star, 
$f_{d1,c}$ to be
\begin{equation}
f_{d1,c} = (1+\gamma_R)^{-3} \approx 3/4.
\end{equation}
However, for planets which orbit the secondary star, the completeness is
\begin{equation}
f_{d2,c} = (1+\gamma_R^{-1})^{-3} \approx 3/4000.
\end{equation}

\begin{comment}
We note in passing that we are considering a SNR-limited survey, which in this 
case is the same as a magnitude-limited one.
The observer might prefer to perform a completeness-limited survey, which in 
this case would be the same as a volume-limited one with limiting distance
$d_{{\rm max},d}^{\rm p}$.
This change might circumvent various biases.
\end{comment}
With our model defined, we ask:

@@

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
Note that these latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
multiple systems that are higher order multiples as doubles.

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
stated in terms of a limiting magnitude $m_{{\rm min}}$.
In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
consequence of imposing a magnitude cut, the maximum distance to which binary 
stars will be selected is greater than that of single stars.
The ratio of double to single systems in the observed SNR-limited sample is 
then
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s}
	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
\label{eq:beta_model1}
\end{align}
In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
SNR-limited sample.
Since the completeness fraction for binary systems is $1/8$, there are 
$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
planets as single star systems in the system that can yield planets.

@@

The observer who has never heard of binary star systems wants to find the 
occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars of 
radius $R_1$ and luminosity $L_1$.
They detected $N_{\rm det}$ transit signals that appear to be from planets of 
this desired type.
The occurrence rate they then find is
\begin{equation}
\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
\end{equation}

@@

For the mass-luminosity relation, for analytic convenience we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
As a simple check, we fit lines to measured masses and luminosities of dwarf 
stars\footnote{
    Data collected by Torres et al. [2009] for dwarfs earlier than 
    spectral class M, and Benedict et al. [2016] for dwarfs later than 
    spectral class M.
    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
    luminosities, we interpolated over E. Mamajek's 
    table
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float. The 
resulting fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L\propto M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final addition to our model is that we must assume a stellar mass-radius 
relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
Since again, our main purpose is to develop intuition about observational 
biases, we coarsely assume
\begin{equation}
R\propto M.
\end{equation}
\begin{comment}
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
\end{comment}
With our model specified, we ask:

@@

The true occurrence rate is the average number of planets of the desired type 
per star of the desired type.
The number of planets of the desired type is
\begin{equation}
N_{\rm planets} = \Gamma_{t,s}N_s + \Gamma_{t,d}N_d + 
\Gamma_{t,d}w_{d2}f_{{\rm desired}}N_d,
\end{equation}
where as before $N_s$ and $N_d$ are the number of single and double systems, 
and the fraction $f_{{\rm desired}}$ is defined so that $f_{{\rm desired}} 
N_d$ is the number of secondaries that are of the desired type, and as before 
$\Gamma_{t,d}w_{d2}$ is the average number of planets each of these desired 
stars hosts.
Explicitly,
\begin{equation}
f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
        ml}(q)\,{\rm d}q }
{\int_{0.1M_1}^{M_1} {\rm prob}_{\rm 
        ml}(q)\,{\rm d}q}.
\end{equation}
For instance, assuming a primary mass of $M_1 = M_\odot$, and defining the 
``desired type'' of stars being to be $0.7<M_\star/M_\odot<1.3$, we find 
$f_{\rm desired}\approx0.475$.

Counting the number of desired stars that are single, $N_s$, is the same as in 
Sec.~\ref{sec:model_1}. However now the maximum selected distance for a binary 
system is now a function of the system's light ratio,
\begin{equation}
d_{\rm max,d} =
d_{\rm max,s}\times (1+\gamma_R)^{1/2},
\end{equation}
where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.
This affects our ability to count the number of desired stars that are double,
since
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3.
\end{equation}

For a given binary system, $q$, $\gamma_R$, and thus $d_{\rm 
	max,d}(\gamma_R)$ are all random variables.
Since $N_d$ is a function of $d_{\rm max,d}$, 
the number of double star systems becomes a random variable, drawn from the 
ensemble of all possible transit surveys.
Transforming the appropriate probability density functions, the distribution 
function for the number of binary systems,
${\rm prob}(N_d)$, can be expressed as 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\label{eq:prob_Nd_raw}
\end{equation}

@@

\begin{figure}[!t]
    \begin{center}
        \includegraphics[width=.5\textwidth]{figures/joint_prob_r_q.pdf}
    \end{center}
    \caption{The joint probability distribution of a binary star's position 
        and 
        mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
        M^{3.5}$. Note that the marginalized 
        mass ratio distribution is uniform at any given distance.}
    \label{fig:joint_prob_r_q}
\end{figure}

\begin{figure}[!t]
    \begin{center}
        \includegraphics[width=0.5\textwidth]{figures/q_distribn_mag_limited.pdf}
    \end{center}
    \caption{The distribution of the mass ratio for a magnitude limited 
        sample of binary stars. The underlying mass ratios are drawn from 
        Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
        volume-limited sample.
        The entire bias can be understood analytically 
        (Eq.~\ref{eq:pdf_q_mag_limited}) to good precision. 
        The numerical comparison uses the ``realistic'' empirical 
        mass-luminosity relation shown in Fig.~\ref{fig:mass_luminosity}.
    }
    \label{fig:q_distribn_mag_limited}
\end{figure}
